---
title: "Simulate messy epi data"
author: "Anders Gonçalves da Silva"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(bsibayes)
library(dplyr)
library(ggplot2)
library(knitr)
library(bayesplot)
```

# Scenario 1

```{r, sim_data_sc1}
## DISEASE PARAMETERS

disease_type = c(
  "NOTIFIABLE",
  "NON-NOTIFIABLE-CONST",
  "NON-NOTIFIABLE-DEC",
  "NON-NOTIFIABLE-INC"
)

pi = c(
  "notifiable" = 0.05,
  "nonnot_1" = 0.02,
  "nonnot_2" = 0.07,
  "nonnot_3" = 0.05
)

beta = c(
  "notifiable" = 0,
  "nonnot_1" = 0,
  "nonnot_2" = -0.01,
  "nonnot_3" = +0.01
)

DISEASE_PARAM = tibble::tibble(
  disease_type = disease_type,
  pi = pi,
  beta = beta
)

DISEASE_PARAM

## INSTITUTION PARAMETERS
lambda = c(
  'small' = 100,
  'medium' = 250,
  'large' = 625
)

n_inst = 10
n_small = 5
n_medium = 3
n_large = 2

n_small_notreporting = 1
n_medium_notreporting = 1
n_large_notreporting = 1

inst_size = c(rep('small', n_small),
                rep('medium', n_medium),
                rep('large', n_large))

gen_report_status <- function(total, n_notreporting) {
  return(c(rep(TRUE, total - n_notreporting), 
                       rep(FALSE, n_notreporting)))
}

inst_report_status = c(gen_report_status(n_small, n_small_notreporting),
                       gen_report_status(n_medium, n_medium_notreporting),
                       gen_report_status(n_large, n_large_notreporting))
                       
INST_PARAM = tibble::tibble(
  id = 1:n_inst,
  size = inst_size,
  lambda = lambda[inst_size],
  reporting = inst_report_status
)

INST_PARAM

total_sims = nrow(INST_PARAM) * nrow(DISEASE_PARAM)
total_sims
# GENERATE SOME SEEDS
set.seed(42)
seeds = runif(n = total_sims, 1,.Machine$integer.max)
seeds

n_years = 10
base_year = 2001
years = seq(base_year, (base_year + n_years - 1))

sims_tab = bsibayes::sim_counts(disease_param = DISEASE_PARAM, 
                      inst_param = INST_PARAM, 
                      years = years, seed = 42)
```

```{r, plot_sims, fig.cap = "Simulated data counts across 10 institutions"}
ggplot(sims_tab$long_sim, aes(x = year, y = count, 
                     colour = disease_type)) +
  geom_line() +
  facet_wrap(~labcode, nrow = 5) +
  xlab("Years") +
  ylab("Observed Count")
```

## Model 1: Recovering the rate of *notifiable* disease

The simplest model we can use here is assume that the observed counts are 
distributed as Poisson distribution, with parameter $\lambda$ and offset 
equal to $log(BSI)$. The parameter $\lambda$ can have a flat Uniform prior
with domain across $Real^+$ numbers. We can represent this model in the following
way:

$$
\begin{aligned}
  \lambda &\sim Uniform(-\infty,\infty) \\
  Count &\sim Poisson(e^{(log(BSI) + \lambda)})
\end{aligned}
$$

Below, we demonstrate how this model is implemented with `STAN` and `R` using only
the data from the notifiable disease. As we can see from
figure \ref{fig:m1_posterior}, the simulated value of $\pi$ as 0.05 is recovered by 
the model. However, to demonstrate this further,
we performed a posterior check by simulating counts from the posterior values, 
and comparing them to the observed counts from the simulated data. As can be 
seen in figure \ref{fig:m1_posterior_check}, our posterior check simulation 
confirms that the model is correctly capturing the information in the data, and
returning good inferences about $\pi$

### Model 1 `STAN` and `R` code

#### Model 1 fitting

```{r, m1_fitting, eval = T}
m1_fit <- run_models_ni(sims_tab$long_sim, model = 'model1')
```

```{r, fig.cap="Posterior distribution of lambda using model 1. \\label{fig:m1_posterior}"}
posterior <- exp(as.matrix(m1_fit, pars='lambda'))
bayesplot::mcmc_combo(posterior)
```

```{r, }
color_scheme_set("purple")
pp <- bsibayes::posterior_predict(m1_fit, draws=500, seed=42)
ppmat <- pp %>% select(labcode, year,rep, count_hat) %>% tidyr::spread(rep, count_hat) %>% select(-labcode, -year) %>% as.matrix()
bayesplot::ppc_intervals_grouped(y = attr(m1_fit, 'data')$count, yrep = t(ppmat), x = attr(m1_fit, 'data')$year, group = attr(m1_fit, 'data')$labcode, prob = 0.5) +
  labs(
    x = "Year",
    y = "Count BSI",
    title = "50% posterior predictive intervals \nvs count of BSI per lab by year"
  ) +
  panel_bg(fill = "gray95", color = NA) +
  grid_lines(color = "white")
```
